### 2020-06-04 writeup/intro.tex
intro

\begin{abstract}
%When does multi-task learning provide positive (or negative) information transfer, compared to single-task learning?
	Multi-task learning is a powerful approach in many applications such as image and text classification.
	Yet, there is little rigorous understanding of when multi-task learning performs well, even compared to single-task learning.
	To be precise, we say that multi-task learning provides \textit{positive transfer} if it outperforms single-task learning, or \textit{negative transfer} otherwise.
	In this work, we consider three components including \textit{task model distance}, \textit{sample size} and \textit{covariate shift} to determine positive vs. negative transfer.
	We formally quantify each component in the setting of high-dimensional linear regression,
% the performance of predicting a particular task given multiple tasks.
%a multi-task learning architecture that uses a shared layer for all tasks and a separate output layer for each task.
	and show that the bias-variance tradeoff of the multi-task estimator determines the effect of transfer..
%	We analyze the bias-variance tradeoff by developing tight concentration results using random matrix theory.
	%we show that the type of transfer is determined by the bias-variance tradeoff of multi-task learning estimators.
	%We develop a technical tool to analyze the.
%	We develop a technical tool on the inverse of the sum of two random matrices that naturally arises from the analysis.
	Then, we show that the performance of multi-task learning can be negatively affected as a result of: (a) increased task model distance;
	(b) increased source sample size; (c) covariate shift under increased source sample size.
	We validate the above results on text classification tasks.
	Inspired by our theory, we show that single-task performance can help understand multi-task performance, and propose an incremental training scheduler for improving the efficiency of multi-task training.% for predicting a particular task.
%	We validate our theory and our proposed method .
\end{abstract}

\section{Introduction}

@@

%\textbf{Setup and Main Results.}
In this work, we apply recent development in the random matrix theory literature to show how to determine positive vs. negative transfer in the setting of high-dimensional linear regression \cite{HMRT19,BLLT20}.
This setting allows us to rigorously quantify three components, defined as follows.
\squishlist
	\item \textbf{Task model distance:} Assume that each task follows a linear model with parameters $\beta_i\in\real^p$, for $1\le i\le t$.
	Hence, the distance between two tasks can be measured by the distance of their model parameters.
	\item \textbf{Sample size:} Let $n_i = \rho_i \cdot p$ denote the sample size of task $i$, for a fixed value $\rho_i > 1$ that does not grow with $p$, for any $1\le i\le t$.
	Hence, for two tasks $i$ and $j$, their data ratio is measured by $n_i / n_j = \rho_i / \rho_j$.
	Importantly, we assume that for a particular (say $t$-th) task, $\rho_t$ is a small constant (say 2), to capture the need for adding more labeled data.
	\item \textbf{Covariate shift:} We assume that for every task $1\le i\le t$, its features are random vectors with covariance matrix $\Sigma_i\in\real^{p\times p}$.
	%$x = \Sigma_i^{1/2}z$, where $z\in\real^p$ consists of i.i.d. entries with mean zero and unit variance, and is a positive semidefinite matrix.
	Given two tasks $i$ and $j$, covariate shift by $\Sigma_i^{1/2}\Sigma_j^{-1/2}$.
\squishend
To motivate our study, Figure \ref{fig_model_shift_phasetrans} shows three examples for two linear regression tasks solved using a shared linear layer for both tasks and an output layer for each task.
We observe that negative transfer occurs as (a) \textit{task similarity}: tasks become more different; (b) \textit{data size}: source/target data size increases.
%\textbf{Task similarity:}
%\textbf{Data sizes:}
%\textbf{Covariate shift:}
Furthermore, MTL performance is negatively affected when (c) \textit{covariate shift}: the covariance matrices of the two tasks become more different.
The observations highlight the need to develop generalization bounds that scale tightly with properties of multiple tasks data.
% that only contains limited amount of labeled data.
%Following Hastie et al.  and Bartlett et al. \cite{},

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.pdf}
		\vspace{-0.075in}
		\caption{Task similarity}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.pdf}
		\vspace{-0.075in}
		\caption{Sample size}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.pdf}
		\vspace{-0.075in}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{%Three takeaways of our theory in Section \ref{sec_insight}.
	We observe that MTL performance
	(a) transitions from positive to negative transfer as \textit{task model distance} increases (Prop.\;\ref{prop_dist_transition});
	(b) transitions from positive to negative transfer as source/target \text{sample size} increases (Prop.\;\ref{prop_data_size});
	(c) worsens as \textit{covariate shift} gets more severe (larger $\lambda$) and source/target \textit{sample size} increases (Prop.\;\ref{prop_covariate}).
	The $y$-axis measures the loss of STL minus MTL.}
	\label{fig_model_shift_phasetrans}
	\vspace{-0.2in}
\end{figure}

>>>>

\vspace{-0.2in}
\begin{abstract}
%When does multi-task learning provide positive (or negative) information transfer, compared to single-task learning?
	Multi-task learning is a powerful approach in many applications such as image and text classification.
	Yet, there is little rigorous understanding of when multi-task learning performs well, even compared to single-task learning.
	To be precise, we say that multi-task learning provides \textit{positive transfer} if it outperforms single-task learning, or \textit{negative transfer} otherwise.
	In this work, we consider three components including \textit{task model distance}, \textit{sample size} and \textit{covariate shift} to determine positive vs. negative transfer.
	We formally quantify each component in the setting of high-dimensional linear regression,
% the performance of predicting a particular task given multiple tasks.
%a multi-task learning architecture that uses a shared layer for all tasks and a separate output layer for each task.
	and show that the bias-variance tradeoff of the multi-task estimator determines the effect of transfer..
%	We analyze the bias-variance tradeoff by developing tight concentration results using random matrix theory.
	%we show that the type of transfer is determined by the bias-variance tradeoff of multi-task learning estimators.
	%We develop a technical tool to analyze the.
%	We develop a technical tool on the inverse of the sum of two random matrices that naturally arises from the analysis.
	Then, we show that the performance of multi-task learning can be negatively affected as a result of: (a) increased task model distance;
	(b) increased source sample size; (c) covariate shift under increased source sample size.
	We validate the above results on text classification tasks.
	Inspired by our theory, we show that single-task performance can help understand multi-task performance, and propose an incremental training scheduler for improving the efficiency of multi-task training.% for predicting a particular task.
%	We validate our theory and our proposed method .
\end{abstract}

\vspace{-0.1in}
\section{Introduction}

@@

%\textbf{Setup and Main Results.}
In this work, we apply recent development in the random matrix theory literature to show how to determine positive vs. negative transfer in the setting of high-dimensional linear regression \cite{HMRT19,BLLT20}.
This setting allows us to rigorously quantify three components, defined as follows.
For simplicity, we consider two tasks with first being the source and the second being the target, but our formulation applies more generally.
\squishlist
	\item \textbf{Task model distance:} Assume that each task follows a linear model with parameters $\beta_i\in\real^p$, for $i=1, 2$.
	Hence, $\norm{\beta_1 - \beta_2}$ measure the distance of the two tasks.
	\item \textbf{Sample size:} Let $n_i = \rho_i \cdot p$ denote the sample size of task $i$, for a fixed value $\rho_i > 1$ that does not grow with $p$, for any $1\le i\le t$.
	Hence, the source/target sample ratio is measured by $n_1 / n_2 = \rho_1 / \rho_2$.
	Importantly, we assume that $\rho_2$ is a small constant (say 2), to capture the need for adding more labeled data.
	\item \textbf{Covariate shift:} We assume that for every task, its features are random vectors with covariance matrix $\Sigma_i\in\real^{p\times p}$.
	%$x = \Sigma_i^{1/2}z$, where $z\in\real^p$ consists of i.i.d. entries with mean zero and unit variance, and is a positive semidefinite matrix.
	We measure covariate shift by $\Sigma_i^{1/2}\Sigma_j^{-1/2}$.
\squishend
To motivate our study, Figure \ref{fig_model_shift_phasetrans} shows three examples for two linear regression tasks solved using a shared linear layer for both tasks and an output layer for each task.
We observe that negative transfer occurs as (a) \textit{task similarity}: tasks become more different; (b) \textit{data size}: source/target data size increases.
%\textbf{Task similarity:}
%\textbf{Data sizes:}
%\textbf{Covariate shift:}
Furthermore, MTL performance is negatively affected when (c) \textit{covariate shift}: the covariance matrices of the two tasks become more different.
The observations highlight the need to develop generalization bounds that scale tightly with properties of multiple tasks data.
% that only contains limited amount of labeled data.
%Following Hastie et al.  and Bartlett et al. \cite{},

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.pdf}
		\vspace{-0.075in}
		\caption{Task similarity}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.pdf}
		\vspace{-0.075in}
		\caption{Sample size}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.pdf}
		\vspace{-0.075in}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{%Three takeaways of our theory in Section \ref{sec_insight}.
	We observe that MTL performance
	(a) transitions from positive to negative transfer as \textit{task model distance} increases (Prop.\;\ref{prop_dist_transition});
	(b) transitions from positive to negative transfer as source/target \text{sample size} increases (Prop.\;\ref{prop_data_size});
	(c) worsens as \textit{covariate shift} gets more severe (larger $\lambda$) and source/target \textit{sample size} increases (Prop.\;\ref{prop_covariate}).
	The $y$-axis measures the loss of STL minus MTL.}
	\label{fig_model_shift_phasetrans}
	\vspace{-0.3in}
\end{figure}
